Evaluating staged investments in critical infrastructure for climate adaptation

### 2019-04-11 slides-IPSWD/presentation.tex
slides for IPSWD

\title{H52F-05H: Robust Adaptation to\\Multi-Scale Climate Variability}
\subtitle{Toward Better Water Planning and Management in an Uncertain World I}
\date{14 December 2018}
\author{\alert{James Doss-Gollin}$^1$, David J. Farnham$^2$, Scott Steinschneider$^3$, Upmanu Lall$^1$}
\institute{
  $^1$Columbia University Department of Earth and Environmental Engineering\\
  $^2$Carnegie Institution for Science\\
  $^3$Department of Biological and Environmental Engineering, Cornell University}
\titlegraphic{\hfill\includegraphics[height=1.25cm]{SeasCrown_blue.png}}

@@

% Package List
\usepackage{
  array,                              % for custom table widths
  appendixnumberbeamer,               % don't count appendix slides in progress bar
  booktabs,                           % for better (alternative?) tables
  natbib,                             % references!
	physics,                            % for better notation
  siunitx,                            % for SI notation
}

@@

% Use the glossaries package
\usepackage[acronym]{glossaries}
\makeglossaries
\newacronym{acc}{ACC}{anthropogenic climate change}
\newacronym{amo}{AMO}{Atlantic Multidecadal Oscillation}
\newacronym{cba}{CBA}{Cost-Benefit Analysis}
\newacronym{enso}{ENSO}{the El Ni\~{n}o-Southern Oscillation}
\newacronym{ffa}{FFA}{flood frequency analysis}
\newacronym{gcm}{GCM}{general circulation model}
\newacronym{hmm}{HMM}{hidden Markov model}
\newacronym{iid}{IID}{independent and identically distributed}
\newacronym{ipcc}{IPCC}{International Panel on Climate Change}
\newacronym{ipo}{IPO}{Interdecadal Pacific Oscillation}
\newacronym{lbda}{LBDA}{living blended drought analysis}
\newacronym{lfv}{LFV}{low-frequency climate variability}
\newacronym{nao}{NAO}{North Atlantic Oscillation}
\newacronym{npv}{NPV}{Net Present Value}
\newacronym{pdo}{PDO}{Pacific Decadal Oscillation}
\newacronym{s2d}{S2D}{seasonal to decadal}
\newacronym{s2s}{S2S}{sub-seasonal to seasonal}

@@

\begin{frame}{Motivating Example}
  What to do after Sandy? \citep{CityofNewYork:2013uh}
  \begin{figure}
    \centering
    \includegraphics[height=0.35\textheight]{three-barriers.png}~
    \includegraphics[height=0.35\textheight]{dull-disasters.png}\\
    \includegraphics[height=0.3\textheight]{flood-concept.png}~
    \includegraphics[height=0.3\textheight]{coastal-resiliency.png}
  \end{figure}
\end{frame}

@@

\begin{frame}{Idea 1: Risk Estimates over Finite Future Periods}
  \begin{alertblock}{Typical Approach:}
    \acrfull{cba}, probably with discounting, over a \alert{finite} planning horizon of $M$ years.
  \end{alertblock}
  \pause
  Project should be evaluated on climate conditions over this finite planning period:
  \begin{itemize}
    \item For ``mega-project'', $M \geq \SI{50}{years}$
    \item For small, flexible project, $M \leq \SI{5}{years}$
  \end{itemize}
\end{frame}

@@

\begin{frame}{Experiment Setup}
  \begin{alertblock}{Research Objective}
    How well can one identify \& predict cyclical and secular climate signals over a finite planning period ($M$),  given limited information?
 \end{alertblock}
  Let $P^* \def \mathbb{P} \qty(X > X^*)$.
  Note that the insurance premium (or risk factor) is:
  \begin{equation*}
    R = \mathbb{E}\qty[ P^*] + \lambda \mathbb{V} \qty[P^*]
  \end{equation*}
  \pause
  \begin{columns}
    \begin{column}{0.5\textwidth}
      \begin{figure}
        \includegraphics[width=\textwidth]{Example-NINO3-M100-N50-Stationary-Trend.pdf}
      \end{figure}
    \end{column}
    \begin{column}{0.5\textwidth}
      Systematic, stylized experiments:
      what happens as we vary $M$, $N$, climate structure, estimating model?
    \end{column}
  \end{columns}
\end{frame}

@@

\begin{frame}{Summary}
  \begin{columns}[T]
    \begin{column}{0.5\textwidth}
      Assertions:
      \begin{itemize}
        \item Investment evaluation depends on climate condition over finite planning period
        \item Physical hydroclimate systems vary on many scales
        \item Physical drivers of risk depend on planning period
      \end{itemize}    
    \end{column}
    \pause
    \begin{column}{0.5\textwidth}
      Implications:
      \begin{itemize}
        \pause
        \item \emph{Ability} to identify and predict different climate signals depends on information available (\eg, $N$)
        \pause
        \item \emph{Importance} of predicting different climate signals depends on extrapolation desired (\ie, planning period)
        \pause
        \item In general, low risk tolerance and/or limited information favor investments with short planning periods.
      \end{itemize}
    \end{column}
  \end{columns}
\end{frame}

@@

\begin{frame}{Example Sequences and Fits}
  \begin{figure}
    \includegraphics[width=\textwidth]{Example-NINO3-M100-N50.pdf}
    \caption{Example of sequences generated with $M=100$ and $N=50$}
  \end{figure}
\end{frame}

>>>>

\title{Evaluating staged investments in critical infrastructure for climate adaptation}
\subtitle{Interdisciplinary Ph.D. Workshop in Sustainable Development 2019}
\date{13 April 2019}
\author{\alert{James Doss-Gollin}$^1$, David J. Farnham$^2$, Scott Steinschneider$^3$, Upmanu Lall$^1$}
\institute{
  $^1$Columbia University Department of Earth and Environmental Engineering\\
  $^2$Carnegie Institution for Science\\
  $^3$Department of Biological and Environmental Engineering, Cornell University}
\titlegraphic{\hfill\includegraphics[height=1.25cm]{SeasCrown_blue.png}}

@@

% Package List
\usepackage{
  amsmath,amssymb,                    % for more math commands
  array,                              % for custom table widths
  appendixnumberbeamer,               % don't count appendix slides in progress bar
  booktabs,                           % for better (alternative?) tables
  natbib,                             % references!
	physics,                            % for better notation
  siunitx,                            % for SI notation
}

@@

% Use the glossaries package
\usepackage[acronym]{glossaries}
\makeglossaries
\newacronym{acc}{ACC}{anthropogenic climate change}
\newacronym{amo}{AMO}{Atlantic Multidecadal Oscillation}
\newacronym{cba}{CBA}{Cost-Benefit Analysis}
\newacronym{enso}{ENSO}{the El Ni\~{n}o-Southern Oscillation}
\newacronym{ffa}{FFA}{flood frequency analysis}
\newacronym{gcm}{GCM}{general circulation model}
\newacronym{hmm}{HMM}{Hidden Markov Model}
\newacronym{iid}{IID}{independent and identically distributed}
\newacronym{ipcc}{IPCC}{International Panel on Climate Change}
\newacronym{ipo}{IPO}{Interdecadal Pacific Oscillation}
\newacronym{lbda}{LBDA}{living blended drought analysis}
\newacronym{lfv}{LFV}{low-frequency climate variability}
\newacronym{nao}{NAO}{North Atlantic Oscillation}
\newacronym{npv}{NPV}{Net Present Value}
\newacronym{pdo}{PDO}{Pacific Decadal Oscillation}
\newacronym{s2d}{S2D}{seasonal to decadal}
\newacronym{s2s}{S2S}{sub-seasonal to seasonal}

@@

\begin{frame}{The really big picture}
  \begin{figure}
    \centering
    \pause
    \includegraphics[height=0.325\textheight]{powerless.png}~
    \includegraphics[height=0.325\textheight]{ho-2017-1.png}\\
    \pause
    \includegraphics[height=0.325\textheight]{jongman-fig-2.jpg}~
    \includegraphics[height=0.325\textheight]{wapo-usa-temp.png}
  \end{figure}
  \nocite{economist-nigeria:2016,Ho:2017gy,Jongman:2012cr}
\end{frame}

\begin{frame}{Motivating case study}
  What to do after Sandy? \citep{CityofNewYork:2013uh}
  \begin{figure}
    \centering
    \includegraphics[height=0.35\textheight]{three-barriers.png}~
    \includegraphics[height=0.35\textheight]{dull-disasters.png}\\
    \includegraphics[height=0.3\textheight]{flood-concept.png}~
    \includegraphics[height=0.3\textheight]{coastal-resiliency.png}
  \end{figure}
\end{frame}

\begin{frame}{Today}
  My perspective: sequential planning under uncertainty
  \pause
  \begin{enumerate}
    \item How does organized variability in the climate inform this planning?
    \pause
    \item Observations $\Rightarrow$ notice three interesting things about the world
    \pause
    \item Stylized computational experiments $\Rightarrow$ understand implications in an idealized system
  \end{enumerate}
  \pause
  Paper submitted to Earth's Future; all codes are available at \url{http://github.com/jdossgollin/2018-robust-adaptation-cyclical-risk}
\end{frame}

@@

\begin{frame}{Idea 1: Risk Estimates over Finite Future Periods}
  \begin{alertblock}{Typical Approach:}
    \acrfull{cba}, probably with discounting, over a \alert{finite} planning horizon of $M$ years.
  \end{alertblock}
  \pause
  Project success depends on climate conditions over this finite planning period:
  \begin{itemize}
    \item For dam, storm barrier: $M \geq \SI{50}{years}$
    \item For cat bond, zoning change: $M \leq \SI{5}{years}$
  \end{itemize}
  \pause
  Small $M$: defer large investment and allow some uncertainties to be resolved
\end{frame}

@@

\begin{frame}{Experiment Setup}
  \begin{alertblock}{Research Objective}
    How well can one identify \& predict cyclical and secular climate signals over a finite planning period ($M$),  given limited information?
 \end{alertblock}
  Let $P^* = \mathbb{P} \qty(X > X^*)$.
  \pause
  We consider
  \begin{itemize}
    \item 3 \alert{idealized scenarios of climate change}
    \item 3 \alert{simple} models for projecting risk
  \end{itemize}
  Measure bias and variance of $P^*$.
  \vfill
  \pause
  \emph{Don't use these models for actual estimation!}
\end{frame}

\begin{frame}{How it works}
  \begin{figure}
    \centering
    \includegraphics[width=\textwidth,height=0.7\textheight]{Example-NINO3-M100-N50.pdf}
  \end{figure}
\end{frame}

@@

\begin{frame}{Summary}
  \begin{columns}[T]
    \begin{column}{0.45\textwidth}
      \begin{itemize}
        \item Investment evaluation depends on climate condition over finite planning period
        \item Physical hydroclimate systems vary on many scales
        \item Physical drivers of risk depend on planning period
      \end{itemize}    
    \end{column}
    \begin{column}{0.55\textwidth}
      \begin{figure}
        \includegraphics[width=\textwidth]{M-N-Sketch.pdf}
      \end{figure}
    \end{column}
  \end{columns}
\end{frame}

\begin{frame}{Conclusions}
  \begin{itemize}
    \item Quasi-periodic and secular climate signals, with different identifiability and predictability, control future uncertainty and risk
    \pause
    \item Adaptation strategies need to consider how uncertainties in risk projections influence success of decision pathways
    \pause
    \item Stylized experiments reveal how bias and variance of climate risk projections influence risk mitigation over a finite planning period
  \end{itemize}
\end{frame}

\section{Next steps}

@@
